---
title: "L50"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---

# SET-UP

```{r markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6,warning = FALSE, message = FALSE)
options(width = 300)
```    

```{r settings, message = F}
source('0.0_settings.R')
```

# Input data
## Read
```{r read, message = F}
new <- FALSE
if(new){
    bio <- get.bio(species='maquereau',user=imlp.user,password=imlp.pass)
    f <- paste0('Rdata/bio_',Sys.Date(),'.Rdata')
    save(bio,file=f)
}else{
    df <- file.info(list.files("Rdata/", full.names = T,pattern="bio_"))
    f <- rownames(df)[which.max(df$mtime)]
    load(f)
}
print(f)
```

## Clean
```{r clean, message = F}
table(bio$matur)

# subset
bio.l50 <- bio[,c('year','nafo','gear','sex','month','length.frozen','agef','matur')]
names(bio.l50)[c(6:7)] <- c('length','age')
nrow(bio.l50)

# what months?
# bio.l50 <- bio.l50[bio.l50$month %in% c(6:7),]   # subset later

# remove NAs
bio.l50 <- bio.l50[!is.na(bio.l50$matur) & 
                   !is.na(bio.l50$length) & 
                   !is.na(bio.l50$year) &
                   !is.na(bio.l50$age),]
nrow(bio.l50)

# remove fish with length/ages (cohorts) that cannot be trusted 
bio.l50 <- bio.l50[!(bio.l50$age>=18&!is.na(bio.l50$age)),]      
bio.l50 <- ddply(bio.l50,c('age'),transform,outlier=outlier(length,coef=3))  
bio.l50[(bio.l50$age==0 & bio.l50$length>300 &!is.na(bio.l50$age)),'outlier'] <- TRUE     
bio.l50[is.na(bio.l50$outlier),'outlier'] <- FALSE
bio.l50 <- bio.l50[bio.l50$outlier==FALSE,]
bio.l50$outlier <- NULL
nrow(bio.l50)

### transform
# cohort
bio.l50$cohort <- with(bio.l50,year-age)
bio.l50 <- bio.l50[bio.l50$cohort>=min(bio.l50$year),]
bio.l50 <- bio.l50[bio.l50$cohort<max(bio.l50$year)-2,]

# mature vs immature
bio.l50$mat <- ifelse(bio.l50$matur<3,0,1) # based on maturity stage, not sex (F, I, M)

# correct maturity stage of age 0 fish. If caught in months 8-12 can impossibly be mature because only some months old (see also maturity-at-age).
bio.l50[bio.l50$mat==1 & bio.l50$age==0 & bio.l50$month>7 &!is.na(bio.l50$age) & !is.na(bio.l50$month),'mat'] <- 0

# length in mm
bio.l50$length <-bio.l50$length*1000

```

## Tables {.tabset}
### n (all)
```{r table n_all}
kable(t(table(bio.l50$mat,bio.l50$cohort)))
```

### n (may-july)s
```{r table n_mayjune}
kable(with(bio.l50[bio.l50$month %in% 5:7,],t(table(mat,cohort))))
```

## Plots {.tabset}

### length-mat (all)
```{r plot length_mat_alldat,fig.width=18}
prop.mat <- ddply(bio.l50,c('cohort','length'),summarise,
               prop.immat=length(mat[mat==0])/length(mat),
               prop.mat=length(mat[mat==1])/length(mat),
               n=length(mat))

ggplot(prop.mat,aes(x=length,y=prop.mat))+
    geom_line(size=1)+
    geom_point(aes(col=n))+
    facet_wrap(~cohort)+
    scale_color_viridis_c()+
    scale_x_continuous(limits=c(100,450),expand = c(0,0))
```

### length-mat (may-jul)
```{r plot mat_gsi,fig.width=18}
prop.mat <- ddply(bio.l50[bio.l50$month %in% 5:7,],c('cohort','length'),summarise,
               prop.immat=length(mat[mat==0])/length(mat),
               prop.mat=length(mat[mat==1])/length(mat),
               n=length(mat))

ggplot(prop.mat,aes(x=length,y=prop.mat))+
    geom_line(size=1)+
    geom_point(aes(col=n))+
    facet_wrap(~cohort)+
    scale_color_viridis_c()+
    scale_x_continuous(limits=c(100,420),expand = c(0,0))
```

### outliers (1) ?
```{r plot length_mat_outliercheck,fig.width=18}
ggplot(bio.l50[bio.l50$month %in% 5:7,],aes(x=mat,y=length,group=mat))+
    geom_point(col='red')+
    geom_boxplot()+
    coord_flip()+
    labs(y='Length (mm)',x='Maturity')
```

### outliers (2)?
Some stage 7 and 8 maturities might actually be immature fish? Even if fully spend, for some fish GSI remains higher that for immature fish. Unsure for mackerel.
```{r plot length_mat_mayjul,fig.width=18}
df <- bio[bio$month %in% 5:7 & bio$agef<3,]

ggplot(df,aes(x=matur,y=wgonad/weight*100,group=matur))+
    geom_boxplot()+
    geom_point(col='grey')+
    scale_y_continuous(limits = c(0,50),expand = c(0,0))
```

# Calculations
## Overall
```{r calcs_tot, message = F}
# exclude 1973 (idem maturity-at-age: something very wrong with this year)
bio.l50 <- bio.l50[bio.l50$year!=1973,]
prop.mat <- prop.mat[prop.mat$year!=1973,]

# select data
d <- bio.l50[bio.l50$month %in% 5:7,]

# functions
get.prop <- function(cf,p=0.5){
  (log(p/(1-p))-cf[1])/cf[2]
}

get.l50 <- function(x,p=0.5){
    cf <- coef(x)
    l <- get.prop(cf,p=p)
    names(l) <- 'l50'
    return(l)
}

# fit (overall: mostly method testing)
# mod <- glm(mat~length,data=d,family = binomial(link=logit))
# set.seed(123)
# boot <- Boot(mod, R=999)
# bootl <- apply(boot$t,1,get.prop)
# bootq <- quantile(bootl,c(0.025,0.975))
# ret <- c(get.l50(mod),coef(mod))
# 
# 
# plot.new()
# plot(mat~length,data=d,xlim=c(0,max(bio.l50$length,na.rm = TRUE)))
# l <- seq(min(bio.l50$length),max(bio.l50$length),1)
# m <-predict(mod,list(length=l),type='response')
# lines(l, m,col='red',lwd=3)
# abline(h=0.5,col='grey',lty=3)
# abline(v=ret[1],col='grey',lty=3)
```

## Annual
```{r calcs_annual, message = F}
if(new){
        # data frame to hold predictions
    preds <- expand.grid(cohort=unique(d$cohort),length=seq(min(d$length),max(d$length),1),pred=NA)
    
    # fit annual model
    l50 <- ddply(d,c('cohort'),function(x){
        co <- unique(x$cohort)
        tab <- c(mat=nrow(x[x$mat==1,]),immat=nrow(x[x$mat==0,]),tot=nrow(x))
        mod <- suppressWarnings(try(glm(mat~length,data=x,family = binomial(link=logit)), silent=TRUE))
        if ('try-error' %in% class(mod)){ 
            ret <- c(l50=NA,low=NA,high=NA,'(Intercept)'=NA,length=NA)
        }else{
          set.seed(123)
          boot <- suppressWarnings(try(Boot(mod, R=999), silent=TRUE))
            if (!'try-error' %in% class(mod)){ 
              bootl <- apply(boot$t,1,get.prop)
              bootq <- quantile(bootl,c(0.025,0.975)) 
              ret <- c(get.l50(mod),low=bootq[1],high=bootq[2],coef(mod))
            }else{
              ret <- c(get.l50(mod),low=NA,high=NA,coef(mod))
            }
            preds[preds$cohort==co,'pred'] <<- predict(mod,list(length=preds[preds$cohort==co,'length']),type='response')
        }
        return(c(ret,tab))
    })
    
    # remove poor fits
    this.l50 <- l50[l50$immat>=10,]               # insufficient data
    this.l50 <- this.l50[this.l50$cohort!=1990,]  # poor model fit (poor data)
    names(this.l50)[3:4] <- c('low','high')
    
    # save outputs
    f <- paste0('Rdata/l50_',Sys.Date(),'.Rdata')
    save(this.l50,file=f)
    
    f <- paste0('Rdata/l50preds_',Sys.Date(),'.Rdata')
    save(preds,file=f)
    
    f <- paste0('csv/l50_',Sys.Date(),'.csv')
    this.l50$l50 <- round(this.l50$l50,0)
    this.l50$low <- round(this.l50$low,0)
    this.l50$high <- round(this.l50$high,0)
    write.csv(this.l50,file=f,row.names = FALSE)
}else{
    dfn <- file.info(list.files("Rdata/", full.names = T,pattern="l50_"))
    load(rownames(dfn)[which.max(dfn$mtime)])
    
    dfn <- file.info(list.files("Rdata/", full.names = T,pattern="l50preds_"))
    load(rownames(dfn)[which.max(dfn$mtime)])
    
    dfn <- file.info(list.files("csv/", full.names = T,pattern="l50_"))
    this.l50 <- read.csv(rownames(dfn)[which.max(dfn$mtime)])
    print(rownames(dfn)[which.max(dfn$mtime)])
}


# info
mean(this.l50$l50)
range(this.l50$l50)
tail(this.l50,4)
```

## Plots {.tabset}

### final
```{r l50_final, message = F,fig.width=10,fig.height=5}
from <- max(this.l50$cohort)
to <- from-10
mu <- mean(this.l50[this.l50$cohort %in% from:to,'l50'])
seg <- data.frame(x1=from,x2=to,y1=mu,y2=mu)
g <- ggplot(this.l50,aes(x=cohort,y=l50))+
  geom_errorbar(aes(ymin=low,ymax=high),width=0)+
  geom_point()+
  geom_text(aes(label=tot,y=Inf),vjust=3,size=1.6)+
  geom_text(aes(label=immat,y=Inf),vjust=6,size=1.6)+
  geom_text(aes(label='  n (total)',y=Inf,x=-Inf),vjust=3,size=1.6,hjust=0)+
  geom_text(aes(label='  n (immature)',y=Inf,x=-Inf),vjust=6,size=1.6,hjust=0)+
  geom_hline(yintercept=268,col='red',linetype='dashed')+
  #geom_text(aes(y=268,x=Inf,label='Current minimum size (268mm)'),col='red',hjust=1,vjust=-0.1)+
  #geom_segment(data=seg,aes(x=x1,xend=x2,y=y1,yend=y2),col='orange',linetype='dashed')+
  #geom_text(aes(y=mu,x=Inf,label='Recent L50 (263mm)'),col='orange',hjust=1,vjust=1.1)+
  scale_y_continuous(limits=c(220,320),expand=c(0,0))+
  scale_x_continuous(expand=c(0,0),limits=c(1970,max(this.l50$cohort)+2))+ 
  labs(y="L50 (mm)",x='Cohort')
g

f <- paste0('img/l50_',Sys.Date(),'.png')
ggsave(f,g,'png',width = 20,height = 8,units = 'cm')
```

### fits
```{r l50_fits, message = F,fig.width=10,fig.height=8}
preds$included <- FALSE
preds[preds$cohort %in% this.l50$cohort,'included'] <- TRUE
ggplot(data=d,aes(x=length,y=mat))+
    geom_point()+
    geom_line(data=preds,aes(y=pred,col=included))+
    facet_wrap(~cohort)+
    labs(y='Maturity',x='Length (mm)')+
    scale_y_continuous(limits=c(0,1),expand=c(0,0),breaks = c(0,1))
```

### Comparison
Values by cohort (not annual) have been used by Smith (terminal year 2020) and Gregoire (terminal year 2007). Assessments in between presented annual L50.
```{r l50_comparison, message = F,fig.width=12}
ggplot(this.l50,aes(x=cohort,y=l50))+
  geom_errorbar(aes(ymin=low,ymax=high))+
  geom_point()+
  geom_text(aes(label=tot,y=Inf),vjust=3,size=2)+
  geom_text(aes(label=immat,y=Inf),vjust=6,size=2)+
  geom_segment(aes(x=1974,xend=2018,y=262,yend=262),col='darkred')+ # values from Andrew 2020 (annual)
  geom_segment(aes(x=2014,xend=2018,y=266,yend=266),col='red')+
  geom_segment(aes(x=1974,xend=2018,y=267,yend=267),col='darkgreen')+ # values from Andrew 2018  (annual)
  geom_segment(aes(x=2013,xend=2018,y=268,yend=268),col='green')+
  geom_segment(aes(x=1973,xend=1979,y=273,yend=273),col='blue')+ # values from Thomas 2016  (annual)
  geom_segment(aes(x=1980,xend=1989,y=260,yend=260),col='blue1')+
  geom_segment(aes(x=1990,xend=1999,y=266,yend=266),col='blue2')+
  geom_segment(aes(x=2000,xend=2009,y=245,yend=245),col='blue3')+
  geom_segment(aes(x=2010,xend=2019,y=266,yend=266),col='blue4')+
  scale_y_continuous(limits=c(210,350),expand=c(0,0))
```
